---
title: "Week 10 Day 2 homework"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(ggfortify)
library(fastDummies)
```

```{r}
diamonds <- read_csv("data/diamonds.csv")

glimpse(diamonds)
```

```{r}
summary(diamonds)
```

```{r}
diamonds %>% 
  ggpairs(columns = c("carat", "x", "y", "z"),
          progress = FALSE)
```

```{r}
diamonds_trim <- diamonds %>% 
  select(- c("X1", "x", "y", "z"))
```

```{r}
diamonds_trim %>% 
  ggpairs(progress = FALSE)
```

```{r}
diamonds_trim %>% 
  ggplot() +
  aes(x = price) +
  geom_histogram() +
  facet_wrap(~ color)
```

```{r}
diamonds_trim %>% 
  ggplot() +
  aes(x = price) +
  geom_histogram() +
  facet_wrap(~ clarity)
```

```{r}
diamonds_trim %>% 
  distinct(cut)


```

Five possible categories => four dummy variables

```{r}
diamonds_trim %>% 
  distinct(color)
```

Seven possible categories => six dummy variables

```{r}
diamonds_trim %>% 
  distinct(clarity)
```

Eight possible catgories => seven dummy variables

```{r}
diamonds_dummy <- diamonds_trim %>% 
  dummy_cols(
    select_columns = c("cut", "clarity", "color"),
    remove_first_dummy = TRUE,
    remove_selected_columns = TRUE
  )
```

```{r}
model_1 <- lm(price ~ carat,
              data = diamonds_dummy)
```

```{r}
autoplot(model_1)
```

```{r}
summary(model_1)
```

Comments: 
Residuals v Fitted looks broadly ok, although there are outliers
QQ plot: Observations not falling on the line at either end, suggesting residuals 
aren't normally distributed
Scale location: Similar to reiduals vs fitted: broadly ok although with outliers
p-value suggests statistically significant
R-squared of 0.85 shows most of variance in price explained by carat
Price ranges between 326 and 18823, so residual error of 1549 is perhaps a bit large.

```{r}
model_2 <- lm(price ~ carat + clarity,
              data = diamonds)
summary(model_2)
```

```{r}
autoplot(model_2)
```

Comments:
- p-values show that the coefficients are statistically significant
- R-squared has increased to 0.89
- residual error has reduced to 1249
However:
- Residuals v fitted shows a shape which suggests the residuals may not be independent
- Scale-Location looks to show increasing residuals
- QQ is again not a straight line suggesting residuals aren't normally distributed


EXTENSION

```{r}
model_3 <- lm(price ~ carat + clarity + log(carat):clarity,
              data = diamonds)
summary(model_3)
```

```{r}
autoplot(model_3)
```

```{r}
library(mosaic)
plotModel(model_3)
```

