---
title: "Week 10 Day 1 homework"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(janitor)
library(broom)
library(ggfortify)
```

```{r}
project <- read_csv("data/project_management.csv")
```

```{r}
project %>% 
  ggplot() +
  aes(x = estimated_length,
      y = actual_length) +
  geom_point()
```

```{r}
project %>% 
  summarise(correlation = cor(estimated_length, actual_length))
```

Actual and estimated lengths are fairly strongly posisitively correlated, so as estimated length increases we expect actual length to increase.

```{r}
model_1 <- lm(actual_length ~ estimated_length, data = project)
model_1
```

```{r}
glance_output <- clean_names(glance(model_1))
glance_output
```

The model can be interpreted as saying that actual length takes 1.2x estimated length + 1.4.
The r-squared says that 0.64 of the variation in actual length can be explained by variation in estimated length.

```{r}
autoplot(model_1)
```

The p-value of 0 means that we can reject the null hypothesis (i.e. that the coefficient = 0), so can conclude that the relationship is statistically significant.

The diagnostic plots do not suggest that the p-value is unreliable.



EXTENSION

```{r}
project %>% 
  ggplot() +
  aes(x = estimated_length,
      y = actual_length) +
  geom_point() +
  geom_text(aes(label = 1:nrow(project)),
            hjust = 0, nudge_x = 0.1)
```

```{r}
autoplot(model_1, which = 4)
```

```{r}
#remove non-influential point
project_2 <- project %>% 
  filter(X1 != 33)

model_2 <- lm(actual_length ~ estimated_length, data = project_2)
model_2
```

These are very similar to the coefficients produced in the first model, so this supports classifying the omitted data point as non-influential.

```{r}
project %>% 
  ggplot() +
  aes(x = estimated_length, y= actual_length) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  geom_smooth(method = "lm", se = FALSE, colour = "blue", data = project_2) 
```



```{r}
#remove influential point
project_3 <- project %>% 
  filter(X1 != 5)

model_3 <- lm(actual_length ~ estimated_length, data = project_3)
model_3
```
These coefficients are differnt from those in the first model. This supports classifying the omitted data point as influential.


```{r}
project %>% 
  ggplot() +
  aes(x = estimated_length, y= actual_length) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  geom_smooth(method = "lm", se = FALSE, colour = "blue", data = project_3) 
```

