---
title: "Week 5 Day 2 Homework"
output: html_notebook
---

```{r}
# Load libraries
library(tidyverse)
library(CodeClanData)
library(lubridate)
library(scales)
```

PART 1
1. Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
qb_revenue_breakdown %>% 
  ggplot() +
  aes(x = Year,
      y = Revenue,
      fill = Product) +
  geom_bar(position = "stack",
           stat = "identity") +
  labs(title = "Revenue by product")
```

2. Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
qb_monthly_sales %>% 
  filter(year(Date) == 2019) %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = Date,
      y = Amount) +
  geom_line() +
  labs(x = "\nMonth",
       y = "Units sold",
       title = "Kwikbit sales - 2019")


```

3. Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
qb_monthly_sales %>% 
  filter(year(Date) == 2019) %>% 
  filter(Cashflow %in% c("Total Revenue", "Cost")) %>%
  ggplot() +
  aes(x = Date,
      y = Amount,
      group = Cashflow,
      colour = Cashflow) +
  geom_line() +
  scale_x_date(breaks = "1 month",
               date_labels = "%b") +
  labs(x = "Month",
       title = "Monthly revenue and costs - 2019")
```
4. Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
qb_competitors %>% 
  ggplot() +
  aes(x = Year,
      y = Revenue/10^6,
      group = Company,
      fill = Company) +
  geom_area() +
  labs(title = "Annual fitness tracker sales",
       y = "Revenue, £m")

```

5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.
```{r}
qb_competitors %>% 
  ggplot() +
  aes(x = Year,
      y = Revenue/10^6,
      group = Company,
      colour = Company) +
  geom_line() +
  geom_point() +
  labs(title = "Annual fitness tracker sales",
       y = "Revenue, £m")
```

6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.
```{r}
qb_device_data %>% 
  mutate(time = (100 * hours) + mins) %>% 
  ggplot() +
  aes(x = time,
      y = counts) +
  geom_line() +
  facet_grid(device ~ id) +
  scale_y_continuous(breaks = c(0, 5000, 10000)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) + 
  labs(x = "Time of day",
       y = "Number of steps",
       title = "Steps recorded") 
  


```

