---
title: "Week 11 Day 3 homework"
output: html_notebook
---

```{r}
library(cluster)
library(dendextend)
library(factoextra)
library(tidyverse)
library(janitor)
library(broom)
```

```{r}
customer_data <- read_csv("data/mall_customers.csv") %>% 
  clean_names() %>% 
  column_to_rownames("customer_id") %>% 
  rename(annual_income = annual_income_k,
         spending_score = spending_score_1_100)


```

```{r}
customer_data %>% 
  ggplot() +
  aes(x = age, y = annual_income, colour = gender) +
  geom_point(size = 2) +
  facet_grid(~ gender)
```

Comment: 
For females, looks like there are groupings
 - between ages 25-40 with annual income between 50k and 75k
 - between 20 and 35 with annual income less than 50k
 For males, looks like there are groupings
  - between 20 and 40 with low annual income
  - between 25 and 50 with annual income c75k

```{r}
customer_data %>% 
  ggplot() +
  aes(x = annual_income, y = spending_score, colour = gender) +
  geom_point(size = 2) +
  facet_grid(~ gender)

```
Comment:
For females, looks like there are five groupings radiating out from a central 
group where nnual_income c50k and spending score c50
 Other four groups then become low income/low score, low income/high score, 
 high income/low score, high income/high score

Similar pattern for males although perhaps not as distinct for the lower income groups


```{r}
customer_data %>% 
  ggplot() +
  aes(x = age, y = spending_score, colour = gender) +
  geom_point(size = 2) +
  facet_grid(~ gender)
```

Comment:
For both females and males looks like there is a grouping of under 40s with high spending score
