---
title: "Week 11 Day 3 homework"
output: html_notebook
---

```{r}
library(cluster)
library(dendextend)
library(factoextra)
library(tidyverse)
library(janitor)
library(broom)
```

```{r}
customer_data <- read_csv("data/mall_customers.csv") %>% 
  clean_names() %>% 
  column_to_rownames("customer_id") %>% 
  rename(annual_income = annual_income_k,
         spending_score = spending_score_1_100)


```

```{r}
customer_data %>% 
  ggplot() +
  aes(x = age, y = annual_income, colour = gender) +
  geom_point(size = 2) +
  facet_grid(~ gender)
```

Comment: 
For females, looks like there are groupings
 - between ages 25-40 with annual income between 50k and 75k
 - between 20 and 35 with annual income less than 50k
 For males, looks like there are groupings
  - between 20 and 40 with low annual income
  - between 25 and 50 with annual income c75k

```{r}
customer_data %>% 
  ggplot() +
  aes(x = annual_income, y = spending_score, colour = gender) +
  geom_point(size = 2) +
  facet_grid(~ gender)

```
Comment:
For females, looks like there are five groupings radiating out from a central 
group where nnual_income c50k and spending score c50
 Other four groups then become low income/low score, low income/high score, 
 high income/low score, high income/high score

Similar pattern for males although perhaps not as distinct for the lower income groups


```{r}
customer_data %>% 
  ggplot() +
  aes(x = age, y = spending_score, colour = gender) +
  geom_point(size = 2) +
  facet_grid(~ gender)
```

Comment:
For both females and males looks like there is a grouping of under 40s with high spending score


```{r}
customer_income_spend <- customer_data %>% 
  select(annual_income, spending_score)
```


```{r}
fviz_nbclust(customer_income_spend, 
             kmeans, 
             method = "wss", 
             nstart = 25)
```

Looks like the "elbow' occurs at k = 5

```{r}
fviz_nbclust(customer_income_spend, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```
Again, maximum occurs at k = 5


```{r}
fviz_nbclust(customer_income_spend, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             k.max = 10)
```
Again, looks like optimal number of clusters is 5


```{r}
clusters <- kmeans(customer_income_spend, centers = 5, nstart = 25)
```

```{r}
clustered_customer <- augment(clusters, customer_data)
```

```{r}
clustered_customer %>% 
  ggplot() +
  aes(x = annual_income, y = spending_score, colour = .cluster) +
  geom_point(size = 2) +
  labs(title = "Clustered customer data",
       x = "Annual income (k)",
       y = "Spending score",
       colour = "Cluster")
```
The clustering seems to be a good fit for this data and confirms the comments made in the initial 
data exploration.

The clusters could be described as:

1. Higher income / higher spending score
 - Incomes above 75k and spending scores of c60 and above. This cluster is quite spread out along the income axis, with no upper bound on income.

2. Lower income / lower spending 
 - Incomes of c40k and below, with spending scores of c45 and below. This cluster is tightly spread along the income axis, as there is a lower bound (0) on income
 
 3. Middle income/middle spend
 - This cluster is tightly packed around the 50k income and 50 spending score region.
