---
title: "NHS Borders Cancer Analysis"
output: 
  html_document:
    theme: flatly
    code_folding: hide
---

***

This analysis looks at cancer incidence in the NHS Borders region and 
makes comparisons with data for Scotland as a whole.

Data is sourced from [https://www.opendata.nhs.scot/](https://www.opendata.nhs.scot/)

> ### Summary
>
> * **The number of incidences of cancer increased by around 33% across NHS Borders between
>     1995 and 2019. ** 
> * **The most incidences of cancer occur in the 65 to 79 age group. **
> * **The most common cancers are those of the skin. **
> * **Incidence rates for some cancers are higher in NHS Borders than the rest of Scotland. **




```{r include = FALSE}
# Load libraries
library(janitor)
library(here)
library(tidyverse)
```

```{r include = FALSE}
#read data
incidence_borders <- read_csv(here("data_raw/incidence_hboard.csv")) %>% 
  clean_names() %>% 
  select(-c("cancer_site_icd10code", "sex_qf")) %>% 
  filter(hb == "S08000016")

summary_borders <- read_csv(here("data_raw/incidence_5yrsummary_hboard.csv")) %>% 
  clean_names() %>% 
  select(-c("cancer_site_icd10code", "sex_qf"),
         -c("crude_rate_lower95pc_confidence_interval":"sir_upper95pc_confidence_interval")) %>% 
  filter(hb == "S08000016")

incidence_scotland <- read_csv(here("data_raw/incidence_scotland.csv")) %>% 
  clean_names() %>% 
   select(-c("cancer_site_icd10code", "sex_qf"),
          -c("crude_rate_lower95pc_confidence_interval":"standardised_incidence_ratio_qf")) %>% 
  rename(hb = country)

summary_scotland <- read_csv(here("data_raw/incidence_5yrsummary_scotland.csv")) %>% 
  clean_names() %>% 
  rename(hb = country) %>% 
  select(-c("cancer_site_icd10code", "sex_qf"),
         -c("crude_rate_lower95pc_confidence_interval":"standardised_incidence_ratio_qf")) 
```


```{r include = FALSE}
borders_counts <- incidence_borders %>% 
  select(hb, 
         cancer_site,
         sex,
         year,
         incidences_all_ages)
```

```{r include = FALSE}
scotland_counts <- incidence_scotland %>% 
  select(hb,
         cancer_site,
         sex,
         year,
         incidences_all_ages)
```

```{r include = FALSE}
borders_counts <- borders_counts %>% 
  bind_rows(scotland_counts) %>% 
  mutate(area = case_when(
    hb == "S08000016" ~ "NHS Borders",
    hb == "S92000003" ~ "Scotland"
  ))
```

#### 1. Number of cancer registrations
The number of cancer registrations in NHS Borders has increased between 1995 and 2019.

The plots below show that the pattern of change in NHS Borders is similar to that in 
Scotland as a whole.

```{r echo = FALSE}
borders_counts %>% 
  filter(cancer_site == "All cancer types" & 
           sex == "All") %>% 
  ggplot(aes(x = year,
             y = incidences_all_ages,
             colour = area)) +
  geom_line() +
  geom_point() +
  scale_colour_manual(values = c("NHS Borders" = "forestgreen",
                               "Scotland" = "blue")) +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  labs(x = "Year of diagnosis",
       y = "Number of registrations",
       title = "New cancer registrations",
       subtitle = "All age groups, all sexes") +
  guides(colour = FALSE) +
  facet_wrap(~ area, scales = "free")


```

From the table below it can be seen that the incidence of cancer in NHS Borders and Scotland
increased by c33% between 1995 and 2019.

```{r echo = FALSE}
borders_counts %>% 
  filter(cancer_site == "All cancer types" & 
           sex == "All" & year  %in% c("1995", "2019")) %>% 
  select(area, cancer_site, year, incidences_all_ages)

```




```{r include = FALSE}
summary_counts_borders <- summary_borders %>%
  select(-c("incidence_rate_age_under5":"crude_rate")) %>% 
  pivot_longer(
    cols = c("incidences_age_under5":"incidences_all_ages"),
    names_to = "incidence_age_band",
    values_to = "incidence_count"
  ) %>%
  mutate(age_band = case_when(
    incidence_age_band %in% c(
      "incidences_age_under5",
      "incidences_age5to9",
      "incidences_age10to14",
      "incidences_age15to19"
    ) ~ "under 19",
    incidence_age_band %in% c(
      "incidences_age20to24",
      "incidences_age25to29",
      "incidences_age30to34"
    ) ~ "20 to 34",
    incidence_age_band %in% c(
      "incidences_age35to39",
      "incidences_age40to44",
      "incidences_age45to49"
    ) ~ "35 to 49",
    incidence_age_band %in% c(
      "incidences_age50to54",
      "incidences_age55to59",
      "incidences_age60to64"
    ) ~ "50 to 64",
    incidence_age_band %in% c(
      "incidences_age65to69",
      "incidences_age70to74",
      "incidences_age75to79"
    ) ~ "65 to 79",
    incidence_age_band %in% c(
      "incidences_age80to84",
      "incidences_age85and_over",
      "incidences_age85to89",
      "incidences_age90and_over"
    ) ~ "over 80",
    incidence_age_band == "incidences_all_ages" ~ "All ages"
  ))
  

```



```{r include = FALSE}
summary_counts_scotland <- summary_scotland %>% 
    select(-c("incidence_rate_age_under5":"crude_rate")) %>% 
  pivot_longer(cols = c("incidences_age_under5":"incidences_all_ages"),
               names_to = "incidence_age_band",
               values_to = "incidence_count") %>% 
  mutate(age_band = case_when(
    incidence_age_band %in% c(
      "incidences_age_under5",
      "incidences_age5to9",
      "incidences_age10to14",
      "incidences_age15to19"
    ) ~ "under 19",
    incidence_age_band %in% c(
      "incidences_age20to24",
      "incidences_age25to29",
      "incidences_age30to34"
    ) ~ "20 to 34",
    incidence_age_band %in% c(
      "incidences_age35to39",
      "incidences_age40to44",
      "incidences_age45to49"
    ) ~ "35 to 49",
    incidence_age_band %in% c(
      "incidences_age50to54",
      "incidences_age55to59",
      "incidences_age60to64"
    ) ~ "50 to 64",
    incidence_age_band %in% c(
      "incidences_age65to69",
      "incidences_age70to74",
      "incidences_age75to79"
    ) ~ "65 to 79",
    incidence_age_band %in% c(
      "incidences_age80to84",
      "incidences_age85and_over",
      "incidences_age85to89",
      "incidences_age90and_over"
    ) ~ "over 80",
    incidence_age_band == "incidences_all_ages" ~ "All ages"
  ))
  

```

```{r include = FALSE}
summary_counts_borders <- summary_counts_borders %>% 
  bind_rows(summary_counts_scotland) %>% 
  mutate(area = case_when(
                          hb == "S08000016" ~ "NHS Borders",
                          hb == "S92000003" ~ "Scotland"),
         age_band = factor(age_band, levels = c("under 19",
                                                "20 to 34",
                                                "35 to 49",
                                                "50 to 64",
                                                "65 to 79",
                                                "over 80",
                                                "All ages")
                           )
         )
```

#### 2. Cancer registrations by age band

The plots below show the total number of cancer registrations between 2015 and 2019,
grouped by age band. New diagnoses are heavily skewed towards older ages (50+),
with the most occuring in the 65 to 79 group. 

A visual inspection suggests that the age distribution in NHS Borders is very similar
to Scotland as a whole.


```{r echo = FALSE}
summary_counts_borders %>% 
  filter(cancer_site == "All cancer types" &
           age_band != "All ages" &
           sex == "All") %>% 
  ggplot() +
  aes(x = age_band,
      y = incidence_count,
      fill = area) +
  geom_col() +
  scale_fill_manual(values = c("NHS Borders" = "forestgreen",
                               "Scotland" = "blue")) +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  labs(x = "Age band",
       y = "Number of registrations",
       title = "New cancer registrations by age band, 2015 - 2019") +
  guides(fill = FALSE) +
  facet_wrap(~ area, scales = "free")


```


#### 3. Most common cancers

The plots below show the most common types of cancer between 2015 and 2019. 
Four of these are skin cancers. 
```{r echo = FALSE}
summary_counts_borders %>% 
  filter(sex == "All" & 
           incidence_age_band == "incidences_all_ages" &
           cancer_site != "All cancer types" &
           area == "NHS Borders") %>% 
  slice_max(incidence_count, n = 10) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidence_count),
      y = incidence_count,
      fill = area) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(values = c("NHS Borders" = "forestgreen",
                               "Scotland" = "blue")) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  coord_flip() +
  labs(y = "Number of registrations",
       title = "Registrations by cancer type, 2015 - 2019"
       ) +
  guides(fill = FALSE) +
  facet_grid(~area, scales = "free")
```

```{r echo = FALSE}
summary_counts_borders %>% 
  filter(sex == "All" & 
           incidence_age_band == "incidences_all_ages" &
           cancer_site != "All cancer types" &
           area == "Scotland") %>% 
  slice_max(incidence_count, n = 10) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidence_count),
      y = incidence_count,
      fill = area) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(values = c("NHS Borders" = "forestgreen",
                               "Scotland" = "blue")) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  coord_flip() +
  labs(y = "Number of registrations",
       title = "Number of registrations by cancer type, 2015 - 2019"
       ) +
  guides(fill = FALSE) +
  facet_grid(~area, scales = "free")


  
```


```{r include = FALSE}
top10_rates_borders <- summary_borders %>% 
  select(-c("year":"incidence_rate_age85and_over")) %>% 
  filter(cancer_site  %in% c("Non-melanoma skin cancer",
                             "Basal cell carcinoma of the skin",
                             "Trachea, bronchus and lung",
                             "Colorectal cancer",
                             "Breast",
                             "Squamous cell carcinoma of the skin",
                             "Colon",
                             "Malignant melanoma of the skin",
                             "Non-Hodgkin lymphoma",
                             "Rectum and rectosigmoid junction"
                             )) 
  
```

```{r include = FALSE}
top10_rates_scotland <- summary_scotland %>% 
  select(-c("year":"incidence_rate_age90and_over")) %>% 
  filter(cancer_site  %in% c("Non-melanoma skin cancer",
                             "Basal cell carcinoma of the skin",
                             "Trachea, bronchus and lung",
                             "Colorectal cancer",
                             "Breast",
                             "Squamous cell carcinoma of the skin",
                             "Colon",
                             "Malignant melanoma of the skin",
                             "Non-Hodgkin lymphoma",
                             "Rectum and rectosigmoid junction"
                             )) 
```

```{r include = FALSE}
top10_rates_borders <- top10_rates_borders %>% 
  bind_rows(top10_rates_scotland) %>% 
  mutate(area = case_when(
    hb == "S08000016" ~ "NHS Borders",
    hb == "S92000003" ~ "Scotland"
  ))
```

#### 4. Incidence rates of most common cancers

The plots below show the incidence _rates_ of the most common types of cancer between 2015 and 2019,
comparing rates between NHS Borders and Scotland. 
A visual inspection of the comparison shows that rates in NHS Borders are higher for some of these cancers.
This suggests that further analysis could be carried out to understand the reasons for these differences.
```{r echo = FALSE}
top10_rates_borders %>% 
  filter(sex == "All") %>% 
  ggplot() +
  geom_linerange(aes(x = reorder(cancer_site, crude_rate), 
                     ymin = 0, 
                     ymax = crude_rate, 
                     colour = area), 
                   position = position_dodge(width = 1)) +
  geom_point(aes(x = reorder(cancer_site, crude_rate),
                 y = crude_rate, 
                 colour = area,
                 shape = area),
               position = position_dodge(width = 1)) +
  scale_colour_manual(values = c("NHS Borders" = "forestgreen",
                               "Scotland" = "blue")) +
  theme_bw() +
  theme(axis.title.y = element_blank()) +
  guides(shape = FALSE) +
  labs(y = "Crude incidence rate",
       title = "Cancer incidence rates",
       subtitle = "10 most common cancers NHS Borders, 2015 - 2019",
       colour = "Area") +
    coord_flip()


```







