---
title: "Week 10 weekend homework"
output: html_notebook
---


```{r}
# Load libraries
library(janitor)
library(tidyverse)
library(lubridate)
library(GGally)
library(modelr)
library(ggfortify)
library(infer)
library(leaps)
```

```{r}
# Read in data
avocados <- read_csv("data/avocado.csv") %>% 
  clean_names()
```

```{r}
glimpse(avocados)
```
```{r}
summary(avocados)
```

```{r}
avocados %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```


### Initial thoughts
x1 is row identifier variable and can be dropped
date - is month a possible predictor? Or create a season variable?
total_bags looks to be sum of small + large + x_large; 
- check this with alias() to ensure no non-standard data/drop aliases
total_volume looks to be sum(4046, 4225. 4770, total_bags) - again check with alias() before dropping aliases
type - check how many distinct
region - check how many distinct before deciding what to do with it

```{r}
distinct(avocados, type)
distinct(avocados, region)
```

# 54 distinct regions: going to drop this variable

```{r}
# Identify any aliased variables
alias(lm(average_price ~ ., 
         data = avocados))
```

# No aliased variables? Seems odd when total_bags, total_volume appear to be made up of other variables

```{r}
avocados %>% 
  mutate(bags = total_bags - small_bags - large_bags  - x_large_bags,
         vol = total_volume - x4046 - x4225 - x4770 - total_bags) %>% 
 filter(bags != 0)
```
Looks like problem is in rounding of number of decimal places. Just going to drop small/large/x_large, x4046, x4225, x4770 variables.
I experimented with rounding to 2dp but it still gave the same problem! Will ask about this on Monday.

```{r}
avocados_trim <- avocados %>% 
  # Make "quarter" variable based on month
  mutate(
    quarter = as.factor(
      case_when(
        month(date) %in% 1:3 ~ "Q1",
        month(date) %in% 4:6 ~ "Q2",
        month(date) %in% 7:9 ~ "Q3",
        month(date) %in% 10:12 ~ "Q4"
      )
    ),
    year = as.factor(year)
    ) %>% 
  select(-c("x1", "date", "x4046", "x4225", "x4770","small_bags", "large_bags", "x_large_bags", "region"))

alias(lm(average_price ~ .,
         data = avocados_trim))
```

