---
title: "Homework - Week 5 Day 1"
output: html_notebook
---

Load libraries
```{r}
library(tidyverse)
library(CodeClanData)
```
**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions.  

```{r}
glimpse(backpack)

```

 

**Question 2.**

The bar chart shows the count of backpacks by weight.

```{r}
backpack %>% 
  ggplot() +
  aes(x = backpack_weight) +
  geom_bar() +
  labs(x = "Backpack weight",
       y = "Count",
       title = "Distribution of backpack weights"
       )
```


**Question 3.**  

The plot below again shows the distribution of backpack weights, but this time shows the distribution for female/male separately.
```{r}
backpack %>% 
  ggplot() +
  aes(x = backpack_weight, 
      group = sex,
      fill = sex) +
  geom_bar(position = "dodge") +
  labs(x = "Backpack weight",
       y = "Count",
       title = "Distribution of backpack weights",
       subtitle = "Split by sex"
       ) +
  scale_x_continuous(breaks = c(5,10,15,20,25,30,35)) +
  scale_y_continuous(breaks = c(5,10))
```


**Question 4.**  

The output below plots body weight (x axis) against backpack weight (y axis). A straight line fitted through the data shows a weak relationship, with backpack weight increasing with bodyweight

```{r}
backpack %>% 
  ggplot() +
  aes(x = body_weight,
      y = backpack_weight) +
  geom_point() +
  geom_smooth(
    aes(x = body_weight,
        y = backpack_weight),
    method = "lm",
    se = FALSE
  ) +
  labs(x = "Body Weight",
       y = "Backpack weight",
       title = "Plot of Body Weight vs Backpack Weight")
```



**Question 5.**  

The plot above is repeated, but this time split by female/male. The fitted line is steeper for females than males, suggesting backpack weight increases faster with bodyweight for females than males. 

```{r}
backpack %>% 
  ggplot() +
  aes(x = body_weight,
      y = backpack_weight,
      group = sex) +
  geom_point() +
  geom_smooth(
    aes(x = body_weight,
        y = backpack_weight),
    method = "lm",
    se = FALSE
  ) +
  labs(x = "Body Weight",
       y = "Backpack weight",
       title = "Plot of Body Weight vs Backpack Weight") +
  facet_wrap(~sex)
```


**Question 6.**  

What **proportion** of students report back pain?   

*** This isn't correct as it is not the proportion.

```{r}
backpack %>% 
  ggplot() +
  aes(x = back_problems) +
  geom_bar() +
  labs(x = "Back Problems", 
       y = "Count",
       title = "Students reporting back pain")
```



**Question 7.**  

Is this proportion different between men and women? 

```{r}
backpack %>% 
  ggplot() +
  aes(x = back_problems, fill = sex) +
  geom_bar(position = "dodge") +
  labs(x = "Back Problems", 
       y = "Count",
       fill = "Sex",
       title = "Students reporting back pain")


```

**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.  

These plot the indicator "backpain" against (1) backpack_weight (2) body_weight and (3) ratio of backpack to body weight, split by sex.

```{r}
backpack %>% 
  mutate(backpain = ifelse(back_problems == 1, "Back pain", "No back pain")) %>% 
  ggplot() +
  aes(x = backpack_weight) +
  geom_bar() +
  labs(x = "Backpack Weight",
       y = "Count",
       title = "Back pain and backpack weight") +
  facet_grid(sex ~ backpain)
 
```

```{r}
backpack %>% 
  mutate(backpain = ifelse(back_problems == 1, "Back pain", "No back pain")) %>% 
  ggplot() +
  aes(x = body_weight) +
  geom_bar() +
  labs(x = "Body Weight",
       y = "Count",
       title = "Back pain and body weight") +
  facet_grid(sex ~ backpain)
```

```{r}
backpack %>% 
  mutate(backpain = ifelse(back_problems == 1, "Back pain", "No back pain")) %>% 
  ggplot() +
  aes(x = ratio) +
  geom_bar() +
  labs(x = "Ratio of Backpack Weight to Body Weight",
       y = "Count",
       title = "Back pain") +
  facet_grid(sex ~ backpain)
```

```{r}
backpack %>% 
  mutate(backpain = ifelse(back_problems == 1, "Back pain", "No back pain")) %>% 
  ggplot() +
  aes(x = body_weight,
    y = backpack_weight) +
  geom_point() +
  labs(x = "Body Weight",
       y = "Backpack Weight",
       title = "Back pain and backpack weight") +
  facet_grid(sex ~ backpain)
```



# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  


**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?
